---
title: "Viz of least and most compact lakes"
author: "Nicole Smith"
date: '2018-01-22'
output:
  html_document: default
  html_notebook: default
---
```{r setup, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE}
library(cowplot)
library(tidyverse)
library(sf)
library(leaflet)

gdb <- "~/.local/share/LAGOS-GIS/LAGOS_US_GIS_Data_v0.1.gdb"

lakes_poly_file <- 'LAGOS_US_All_Lakes_4ha'
states_file     <- 'STATE'

lakes_poly <- st_read(gdb, lakes_poly_file, stringsAsFactors = FALSE)
lakes_poly <- lakes_poly %>%
  st_zm(drop = TRUE) %>%
  filter(Hectares >= 4) %>%
  mutate(FCode = factor(FCode)) %>%
  mutate(FType = factor(FType)) %>%
  select(Permanent_, GNIS_Name, AreaSqKm, ReachCode, FType, FCode, STATE, lagoslakeid, Shape_Area, Shape_Length)

states <- st_read(gdb, states_file) %>%
  st_simplify(TRUE, 5000)
```

```{r functions, echo=FALSE, warning=FALSE, message=FALSE}
# Create a list of ggplots where the polygons share a common (largest) bbox size
# Adapted from: http://rstudio-pubs-static.s3.amazonaws.com/2852_379274d7c5734f979e106dcf019ec46c.html
stage_plots <- function(data, region = NULL, size = c("maxzoom", "scale")) {
  !is.null(region) || stop("you must supply a region identifier")
  
  if(size == "scale"){
    boxes <- list()
    for (i in 1:nrow(data)) {
      boxes[[i]] <- st_bbox(data[i,])
    }
    x_max_range <- max(as.vector(sapply(boxes, function(x) x['xmax'] - x['xmin'])))
    y_max_range <- max(as.vector(sapply(boxes, function(x) x['ymax'] - x['ymin'])))
    max_dim     <- max(y_max_range, x_max_range)
  }
  
  plots <- list()
  for(i in 1:nrow(data)){
    plot_item <- data[i,] %>% as("Spatial") %>% fortify(region = region)
    plot      <- ggplot(plot_item) + 
                    geom_polygon(aes(x = long, y = lat, group = group), 
                            fill = 'lightsteelblue3', color = 'dodgerblue4') +
                    theme_void() + coord_fixed()
    
    if(size == "scale"){
      centroid <- data[i,] %>% st_centroid() %>% st_coordinates()
      x_limits <- c(centroid[,'X'] + .5 * max_dim, centroid[,'X'] - .5 * max_dim)
      y_limits <- c(centroid[,'Y'] + .5 * max_dim, centroid[,'Y'] - .5 * max_dim)
      plot <- plot +
        xlim(x_limits) + ylim(y_limits)
    }
    plots[[i]] <- plot
  }
  return(plots)
}
```

I'll be using the *isoperimetric quotient*, a common measure of compactness that compares the area of the shape to the area of a circle with the same perimeter. $$\frac{4\pi A}{P^2}$$With this measure, a value of 1 indicates the lake is nearly circular, and lower values represent less compactness.
```{r }
lakes <- lakes_poly %>%
  mutate(compactness = (4 * pi * Shape_Area) / (Shape_Length ** 2)) %>%
  arrange(-compactness) %>%
  mutate(compact_rank = row_number())
```

Prepare a few subsets for plotting.
```{r}
top_100 <- lakes %>%
  top_n(100, AreaSqKm) %>%
  arrange(desc(AreaSqKm))
```

#100 largest lakes in LAGOS (4ha+):
```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width="100%"}
staged1 <- stage_plots(top_100, "lagoslakeid", "scale")
p1      <- cowplot::plot_grid(plotlist = staged1)
```

#100 least compact lakes in LAGOS (4ha+):
```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width="100%", eval=FALSE}
staged2 <- stage_plots(least_compact_100, "lagoslakeid")
p2      <- cowplot::plot_grid(plotlist = staged2)
```

#100 lakes of median compactness in LAGOS (4ha+):
```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width="100%", eval=FALSE}
staged3 <- stage_plots(mid_compact_100, "lagoslakeid")
p3      <- cowplot::plot_grid(plotlist = staged3)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
#Prep the layers for leaflet
proj_string <- "+proj=longlat +datum=WGS84"
L.most_compact_100 <- most_compact_100 %>%
  st_transform(proj_string)
L.least_compact_100 <- least_compact_100 %>%
  st_simplify(dTolerance=200) %>%
  st_transform(proj_string)
L.states <- states %>%
  st_transform(proj_string)
```

#Where are the 100 **most** compact lakes?
```{r, echo=FALSE, warning=FALSE, out.width="100%", eval=FALSE}
popup_template <- "<b>GNIS Name:</b><br/>%s<br/><b>FCode:</b><br/>%s"

(m1 <- leaflet() %>%
  addProviderTiles(providers$Esri.WorldImagery, options=providerTileOptions(opacity=.75)) %>%
  addPolygons(data=L.states, weight = 2, opacity = .25, color="white", fill=NULL) %>%
  addPolygons(data=L.most_compact_100, weight = 2) %>%
  addCircleMarkers(data = st_centroid(L.most_compact_100), radius = 4, fillOpacity = 1, stroke = FALSE,
                    popup = ~sprintf(popup_template, GNIS_Name, FCode))
   )
```

#Where are the 100 **least** compact lakes?
```{r, echo=FALSE, warning=FALSE, out.width="100%", eval=FALSE}
(m2 <- leaflet() %>%
   addProviderTiles(providers$Esri.WorldImagery, options=providerTileOptions(opacity=.75)) %>%
   addPolygons(data=L.states, weight = 2, opacity=.25, color="white", fill=NULL) %>%
   addPolygons(data=L.least_compact_100, weight = 2) %>%
   addCircleMarkers(data = st_centroid(L.least_compact_100), radius = 4, fillOpacity = 1, stroke = FALSE,
                    popup = ~sprintf(popup_template, GNIS_Name, FCode))
   )

```

#Are there compactness patterns across the United States?
```{r, echo=FALSE, warning=FALSE, out.width="100%", eval=FALSE}
lake_points <- st_centroid(lakes) %>%
  mutate(x = st_coordinates(.)[,"X"]) %>%
  mutate(y = st_coordinates(.)[,"Y"])

ggplot() +
  stat_summary_hex(data = lake_points, aes(x, y,  z=compactness), binwidth = c(50000, 50000)) +
  geom_sf(data = states, color = 'grey50', fill = NA, size = .5) +
  scale_fill_gradient2(midpoint=median(lake_points$compactness), name = "compactness") +
  theme_void()

```

#Distribution of compactness.
```{r, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
(hist <- ggplot(data=lakes, aes(x=compactness)) +
   geom_histogram(alpha = .5, fill = "blue") +
   geom_vline(aes(xintercept=median(compactness)))
)
```
